On the properties of discs around accreting brown dwarfs

### 2010-05-20 BD_disc/paper.tex
from th -th
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2277 25464f21-6099-402c-b30a-640f1d40d38b

\begin{table*}
\begin{tabular}{|l|l|}
Input parameter&Values (\# values)\\
\hline
Mass ($M_{\odot}$)&0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07 \& 0.08 (8)\\
Age (Myr)&1 \& 10 (2)\\
Rotation period (days)&0.5 \& 5 (2)\\
Areal coverage (of $\dot{M}$, \%)&1 \& 10 (2)\\
Accretion rate (log$(\frac{\dot{M}}{M_{\odot}} yr^{-1})$)&$-6$, $-7$, $-8$, $-9$, $-10$,
$-11$ \& $-12$ (7)\\
Disc mass ($M_*$)&0.01 (1)\\
Surface density profile & $r^{-1}$ (1) \\
Inclination ($^{\circ}$)&0, 27, 39, 48, 56, 64, 71, 77, 84 \& 90
(10)\\
\end{tabular}
\caption{List of all varied input parameters. Resulting in a total
  number of models of 448 (plus 40 models without radiative transfer
  simulations for the naked BDs) and 4480 SEDs (plus 40 for naked
  BDs). \label{par_space_table}}
\end{table*}

@@

\subsubsection{Inner edge of the dust disc}
\label{inner_edge}

@@

As the inner edge moves its temperature is expected to change, this
has been predicted to lead to a correlation of inner edge position and
IR excess \citep{meyer_1997}. This may act to bias surveys correlating
rotation rates with IR excess to search for evidence of
disc-locking. However, the flux from the inner edge will usually peak
between 2 and 3\,$\mu$m \citep{dullemond_2001}, given that the dust
sublimation temperature peaks at $\approx$1400\,K for canonical
densities. This means that for models where dust is significantly
sublimated the inner edge will {\em always} have a temperature of
$\approx$1400\,K and this correlation of disc position and temperature
will be lost.

Equation \ref{inner_eq} shows that as the rotation period, of the
star, increases the co-rotation radius decreases. This will result in,
initially, shorter period systems having closer and hotter inner edges
than their longer period counterparts. In addition, if the accretion
rate is increased in these systems the incident flux on the inner wall
will increase leading to a rise in the temperature of the inner edge.
At some point, in both cases the dust sublimation temperature may be
reached, leading to a change in radial position of the wall. In
addition, the temperature of the inner edge will then tend to a
maximum temperature, of the dust sublimation temperature. Figure
\ref{lum_sub} shows the total stellar luminosity (accretion luminosity
plus the intrinsic stellar luminosity), in log($L_*/L_{\odot}$)
against the change in inner disc location, $R_{\rm co}-R_{\rm inner}$
in units of AU (\textit{top panel}) and $R_*$ \textit{bottom panel}.
This change in inner edge location is from the initially set
co-rotation radius to the final radius after dust sublimation, if
applicable.

The classification of models into typical and extreme accretors is
based on observations (see discussion in \ref{par_accn}) and also
evidence from the simulations themselves. Figure \ref{delta_radius},
\textit{left panels}, shows that dust sublimation does not become
significant until accretion rates of around \logmdot=$-9$ and
increases dramatically with increasing accretion rate thereafter.
This accretion rate is also, unsurprisingly the accretion rate at
which the underlying photospheric emission is overwhelmed by the
accretion flux, and the photospheric features are lost (see Section
\ref{sed_analysis}). The \textit{right panels} of Figure
\ref{delta_radius} show that dust sublimation occurs only for rotation
periods of 0.5 days for typical accretors. For the extreme accretors
the systems with both rotation periods undergo sublimation but the
effect is more significant in the systems with shorter rotation
periods.  The \textit{right panels} also show that for extreme
accretors the older systems undergo more dust sublimation, whereas for
the typical accretors the older systems undergo little or no dust
sublimation.

The density of material in the disc falls with increasing radius from
the star ($\rho\propto R_*^{\alpha}$).  The dust sublimation
temperature is also density dependent \citep{pollack_1994}.
Therefore, for systems where the inner edge has been eroded
significantly from the co-rotation radius the temperature of the inner
edge will fall as the final radius increases.  In this case a
correlation of inner edge temperature and inner edge radius is
expected with a resulting slope determined, at least in part, by the
radial density distribution of the disc.  In this case the the final
inner edge radius and therefore temperature is no longer strongly
correlated with the rotation rate. Figure \ref{inner_temp} shows the
final inner edge location ($R_{\rm inner}$, $R_*$) against the
temperature of the inner wall ($T_{\rm inner}$, K). The \textit{left
  panel} shows the systems designated as typical accretors and the
\textit{right panel} those with extreme accretion rates.  For
\textit{both panels} the systems with rotation periods of 0.5 and 5
days are plotted in red and blue respectively. Systems where $\Delta
R_{\rm inner} < 1R_*$ are plotted ia crosses and $\Delta R_{\rm inner}
> 1R_*$ as filled circles, with the latter denoting systems with
significant sublimation. Significant sublimation is only found for the
extreme accretors, as shown in Figure \ref{delta_radius}.
Furthermore, for the typical accretors Figure \ref{delta_radius} shows
that only the shorter period objects, in general, experience any disc
erosion at the inner edge.

@@

Figure \ref{inner_temp} shows that for those systems where significant
disc erosion ($\Delta R_{\rm inner} > 1.0 R_*$) has occurred the
resulting temperature of the inner edge is weakly correlated with the
inner edge radius, but, critically, not correlated with rotation rate.
The inner edge temperatures of the remaining systems for the extreme
accretors are slightly anti-correlated with the radius to the inner
edge. For the systems with typical acccretion rates, and longer
periods, Figure \ref{inner_temp} shows there is again a weak
correlation between the radius to, and the temperature of the inner
edge. For the shorter period models with typical accretion rates there
is no clear correlation between the temperature at the inner edge and
radius to this edge. The \textit{left panel} of Figure
\ref{inner_temp} shows that, taken as a whole, our models with typical
accretion rates show a clear correlation between the temperature at
the inner edge and the radius to this boundary. This agrees with the
work of \cite{meyer_1997} where this correlation is found for
there $R_{\rm inner}=$1$-$12$R_*$ and $\dot{M_{\rm acc}}=$10$^{-9}$to
10$^{-5} M_{\odot}yr^{-1}$, for flat disc models.
\cite{meyer_1997} use this correlation, and the derivation of
IR magnitudes, to predict a relationship between IR excess and radius
to the inner wall. Our data indicate that this correlation, for
typically accreting systems will translate into a correlation between
rotation rate and IR excess. This could have important implications
for studies of disc-locking where disc presence is examined as a
function of rotation rates, provide an intrinsic bias. In practice
however, this correlation is weak, an unobservable, in our data due to
inner disc wall shape, inclination and flaring effects. This is
discussed in more detail in Section \ref{photometry}.

For simulations without dust sublimation, and analytic density
distributions, the inner disc edge is characterised solely by the
initial disc density prescription. If one includes vertical
hydrostatic equilibrium the scaleheight of the disc inner edge can
change but the radial distribution of dust will not. Dust sublimation
physics will alter the vertical structure, but crucially will also
alter the radial distribution of dust. This means that dust
sublimation effects will control the shape of the inner disc wall and
possibly change this from a vertical wall. Evidence that real inner
disc boundaries (for CTTS) are not vertical walls comes from a lack of
dependence of the derived IR excess on system inclination
\citep{tannirkulam_2007}. As discussed in Section \ref{dust_edge}
the disc inner hole can be created by several mechanisms, however, for
this study we have only included a treatment of magnetic truncation
(initially) and subsequent dust sublimation. The effects of dust
sublimation have been shown to produce curved inner walls due to the
temperature and density dependence of the dust sublimation temperature
\citep{isella_2005}. The curved boundary in this case is
convex, as density increases towards the midplane of the disc, the
dust sublimation temperature increases, meaning the destruction radius
moves closer to the star. A curved inner boundary is also found to
result from dust sublimation in \cite{tannirkulam_2007}. In
\cite{tannirkulam_2007} two populations of dust grains where
included with different scaleheight distributions. This is essentially
a simple model of grain growth and subsequent midplane settling.
\cite{tannirkulam_2007}, found a concave inner disc boundary, this
is due to the larger grains cooling more efficiently. These larger
grains dominate the dust towards smaller vertical distances from the
midplane, and therefore the increased cooling means that the dust
sublimation temperature is reached at closer radii. The implications
of the inner edge shape for the simulated SEDs and photometry are
discussed in Section \ref{sed_analysis} and \ref{photometry}.

For systems where dust sublimation does not occur, or is minimal the
inner edge of the system will remain as a vertical wall, as prescribed
by our initial density distribution. However, as one increases the
flux levels of a given system, by increasing the accretion rates (or
moving to younger ages) the inner edge will increase in
temperature. As the flux levels increase, in a given model, the dust
at the inner edge will sublimate. The dust in the regions of lowest
density or maximum flux interception will be preferentially
destroyed. This will lead to a reduction in the scaleheight and a
slight concave cavity at the inner wall. The inner edge will then move
towards a convex boundary as the inner edge is forced radially outward
towards regions of lower density with increasing flux levels. This
outward movement will result in the outer, vertical, regions of the
dust disc eroding. Increasing the flux still further will lead to
significant dust sublimation and an outward migration. This subsequent
migration will move the inner edge into regions of lower density and
greater scaleheight.

The transition in shape and position of the inner edge for a given
model as the accretion rate, and therefore flux levels, are increased
is shown in Figures \ref{no_sub_183} to \ref{sub_181}. The model
chosen for illustrative purposes has the following parameters,
$M_*$=0.04$M_{\odot}$, Age=1 Myrs and areal coverage=10\%. The model
shown in Figure \ref{no_sub_183} has $\tau$=5 days and accretion
rate \logmdot$=-12$ (model shown in Figure
\ref{flare_-12_183}), with Figures \ref{sub_169} to \ref{sub_181}
showing models with $\tau$=0.5 days and accretion rates \logmdot=$-12$, $-9$, $-7$
and $-6$. The \textit{right
  panels} of Figures \ref{no_sub_183} to \ref{sub_181} show the
temperature structure of the final converged disc structure. The
\textit{left panels} of Figures \ref{no_sub_183} to \ref{sub_181} show
the final log($\rho$) distribution for regions where dust is present,
scaled to a different minimum density (log($\rho_{\rm
  min}$)$=1^{-13}$) compared to Figures \ref{flare_-12_183} and
\ref{flare_-7_193}. The lower (log) density threshold is used in this
case to better trace the inner edge and denser regions of the disc,
whereas Figure \ref{flare_-12_183} and \ref{flare_-7_193} show the
outer diffuse regions of the disc, or disc limits. Additionally, for
Figures \ref{sub_169} to \ref{sub_181} (the sublimated systems), the
\textit{left panels} also show the initial density, in the same units
as the final density, this is shown as a background grey scale (with
the same maximum and minimum as for the final density) to illustrate
the effects of sublimation.

@@

Figure \ref{no_sub_183} shows the inner disc edge as a vertical wall
with no dust sublimation, the inner wall has not changed location from
the initial disc description. The \textit{right panel} of Figure
\ref{no_sub_183} shows that the corresponding temperature structure is
smooth and interpolation at the inner edge yields a temperature of
$~$900\,K. For the shorter period model the inner edge has undergone
some dust sublimation. The \textit{left panel} of Figure \ref{sub_169}
shows that the inner edge has lost dust slightly, preferentially,
along the midplane. Again for Figure \ref{sub_169} the change in inner
edge location is small, $~0.45R_*$ ($~1.0\times 10^{-3}$AU), at the
sublimation (and inner wall) temperature is $~$1450\,K, for a density
of $~2.5\times 10^{-8}$ ($g/cm^3$) (\textit{right panel} of Figure
\ref{sub_169}).The \textit{left panel} of Figure \ref{sub_169} shows
that for even for a negligible accretion rate of \logmdot=$-12$,
sublimation has begun to erode the inner edge, for the younger and
faster rotating systems.  Here the final density distribution has
moved radially outward as can be seen by comparing the grey and colour
scales in the \textit{left panel} of Figure
\ref{sub_169}. Additionally, the inner edge has become slightly
concave. As the flux levels of the central star increase the
temperature of the inner edge surpasses the sublimation temperature at
an increasing radius, destroying more of the disc and pushing the
inner edge radially outwards. The \textit{left panel} of Figure
\ref{sub_175} shows that for an accretion rate of \logmdot$=-9$,
increased sublimation has pushed the inner wall slightly farther from
the star (when compared to Figure \ref{sub_169}), here, $\Delta
R~0.52R_*$ ($~1.1\times 10^{-3}$AU), and $T_{\rm inner}~1450$. Again
the final density distribution has moved radially outward as can be
seen by comparing the grey and colour scales in the \textit{left
  panel} of Figure \ref{sub_175}. The \textit{right panel} of Figure
\ref{sub_175} shows that the gas interior to the inner edge extending
down to the co-rotation radius, which is not sublimated, will reach
temperatures in excess, or equal to, the inner edge, which in this
model is $~$1400\,K. Such hot gas inner discs interior to the dust
disc have been postulated through SED fitting
\citep{eisner_2005,tannirkulam_2008}. Again, the shape of
the inner wall has changed in Figure \ref{sub_175}, the top and bottom
sections of the disc have eroded and the inner wall is convex. The
model in Figure \ref{sub_175} has an accretion rate of \logmdot=$-9$,
which is the transition point between our two groups of accretors,
typical and extreme. Figure \ref{sub_175} shows that for typical
accretion rates, and faster rotation periods (and younger ages)
significant dust sublimation is expected to occur at the inner edge,
resulting in a curved inner wall at around $~$1400\,K. Increasing the
accretion rate, and therefore flux levels still further leads to
significant erosion of the inner disc.  For the extreme accretors the
dust sublimation at the inner edge now changes the inner wall position
by many multiples of the stellar radius. Figures \ref{sub_179} and
\ref{sub_181} are for systems with accretion rates typical in outburst
or episodic accretion systems, \logmdot$=-7$ and $-6$
respectively. The inner edges for these systems are eroded radially by
7.1 and 36.1 ($R_*$) respectively (0.016 and 0.080 AU, respectively),
visible by comparing the grey and colour scales of Figure
\ref{sub_179} and \ref{sub_181}.  As this occurs the disc inner edge
moves far enough that the density drop (with radius) occurring in the
disc causing the sublimation temperature to fall. Consequently the
disc inner edge becomes cooler as it can not sustain temperatures
above the sublimation temperature.  This is shown in the \textit{right
  panels} of Figures \ref{sub_179} and \ref{sub_181}, with approximate
temperatures at the inner edge of 1370 and 1260\,K. In addition, the
inner edge has become concave and, again due to changes in the
underlying disc density distribution, much larger in scaleheight.

A similar sequence is apparent across the grid for set
masses. However, the balance of the rotation period, and therefore
inner edge location, and age and areal coverage, therefore flux
levels, leads to changes in the accretion rate at which the dust
sublimation starts. However, in almost all cases the dust sublimation
does not become significant until at least \logmdot=$-9$ (as discussed
above).

@@

The resulting converged disc structures, as discussed in Section
\ref{seds} are then used to create simulated SEDs. In this section we
explore features in these derived SEDs and link them back to the
physical structures within the disc discussed in Section
\ref{disc_struct}. For analysis purposes our SEDs will comprise
contributions from four main luminosity sources. Firstly, the
accretion luminosity ($L_{\rm acc}$) and the intrinsic stellar
luminosity ($L_{*}$), will effectively be combined and provide flux at
the shorter wavelength section of the SED. For our parameter space the
temperature of the photospheres has a range of $~$3000\,K$<T_{\rm
  eff}<$1600\,K, this will comprise flux contributions peaking in the
range, 0.97 to 1.8\,$\mu$m ($\lambda_{\rm peak}\propto \frac{1}{T_{\rm
    acc}}$). The accretion spot temperatures, for typical accretors,
will range from $~$340$>T_{\rm acc}>$9400\,K, emitting in the range
0.3 to 8.5\,$\mu$m. For the extreme accretors temperatures which can
reach as high as 53\,000\,K, provide flux to $~$0.05\,$\mu$m. The
second pair of luminosity sources are from constituent parts of the
circumstellar disc. The inner edge, ($L_{\rm wall}$), at temperatures
of $~$370 to $~$1500\,K, or flux contributions from $~$1.85 to
7.7\,$\mu$m \citep[nominally 2 to 3\,$\mu$m][]{dullemond_2001}, and
the outer disc ($L_{\rm disc}$), at temperatures of $~$300\,K or less
contributing flux longward of $~$10\,$\mu$m. Disentangling the
accretion and stellar photospheroc flux becomes difficult as $T_{\rm
  acc} \to T_{\rm eff}$, which for an acretion rate of -9 \logmdot,
occurs at a fractional coverages of 20 and 10\%, for rotation periods
of 5 and 0.5 days (with $M_*=0.04M_{\odot}$ and an age of 1 Myr). As
discussed in Section \ref{seds} the {\sc torus} radiative transfer
code tags the photons before they reach the simulated observer in one
of four ways, stellar and thermal, either direct or scattered. These
components allow us to analyse the contributions from the main flux
contributors discussed, i.e. $L_{\rm acc}$ and $L_*$ equals stellar
scattered or direct, with thermal emission comprising $L_{\rm wall}$
and $L_{\rm disc}$. Figure \ref{disc_eg_173} shows a typically
accreting BDD system with the component photon contributions
highlighted, with $M_*=0.04M_{\odot}$ at 1 Myr with an areal coverage
of 10\%, accretion rate of \logmdot=$-10$ and rotation period of 0.5
days. The solid black line is the total simulated SED, with the
stellar photons shown as blue lines and the thermal as red. The
scattered and direct contributions are then shown as dashed and solid
lines respectively. The horizontal and vertical dotted lines then show
the approximate ranges of the flux contributions from the (from left
to right), $L_{\rm acc}$ (blue), $L_*$ (black), $L_{\rm wall}$ (red)
and $L_{\rm disc}$ (black). The vertical dashed lines at the bottom of
Figure \ref{disc_eg_173} show the approximate ranges of the $V$, $I$,
$J$, and $K$ photometric bands we have adopted.

@@

Figure \ref{disc_eg_173} shows that the photosphere and accretion
luminosities dominate the SED at wavelengths shortward of
$~$1$\,\mu$m. After this, for this model, the SED is dominated by
contributions from the disc, at shorter wavelengths, the inner edge
and then at larger wavelengths the outer disc. As shown in Figure
\ref{disc_eg_173} the intrinsic stellar photosphere and accretion
luminosities overlap, in terms of peak wavelength of emission. This
overlap essentially means that the combined stellar flux ($L_*+L_{\rm
  acc}$) will eventually become dominated by the accretion flux, as
one increases the accretion rate. With the features of the intrinsic
stellar photosphere or atmosphere eventually lost or swamped by the
accretion flux. This will lead to difficulties in using spectral
features to identify BDD systems for wavelengths shorter than $~$1
$\mu$m, and, of course, for wavelengths longer than this the SED
becomes dominated by disc emission. Therefore, once the SED of a BDD
system, at shorter wavelengths, becomes dominated by accretion
luminosity spectral classification of these systems (or at least their
central stars) will become unreliable. As discussed in Section
\ref{inner_edge} observations of the inner edge location separate the
BDD systems into two groups, the typical accretors ($\dot{M}\leq$-9
log$(\frac{\dot{M}}{M_{\odot}} yr^{-1})$) and extreme accretors
($\dot{M}>-9$). This division
between typical and extreme accretors is based on observational
results (see Section \ref{par_accn}) as well as consequence of
sublimation. The typical accretors show limited sublimation at the
disc inner edge leading to a weak correlation in inner edge
temperature with edge radius. Whereas, for the extreme accretors the
inner edge is significantly eroded and the dust sublimation
temperature reached, which can not be exceeded. This effectively means
that the inner edge temperature is weakly correlated with temperature
(with a different characteristic slope to the typical accretors), as
the dust sublimation temperature is density dependent, which is in
turn dependent on radial position (see Figures \ref{delta_radius} and
\ref{inner_temp}).

@@

Figure \ref{acc_dom} shows the affect of increasing the accretion
blackbody flux (for increasing accretion rates) for a BD star,
$M=$0.04$M_{\odot}$, at 1 Myr. Whilst the photons originating from the
star (both from $L_*$ and $L_{\rm acc}$) will be tagged as stellar by
{\sc torus} we can separate these flux contributions simply by
observing the naked star system. The panels in Figure \ref{acc_dom}
show the flux from a naked system, with no treatment of the disc. This
enables us to view the effect of increasing accretion rate on the
photospheric flux in isolation. The accretion rates included in all
panels are \logmdot = $-8$, $-9$ and $-12$ (blue, black and red lines
respectively). The \textit{bottom panels} show the systems with a
rotation period of 5 days and \textit{top panels} for those with a
rotation period of 0.5 days. Given our assumption that accretion
occurs from the co-rotation radius, decreasing the rotational period
moves this accretion radius closer to the star, $R_{\rm inner}\propto
\tau^{2/3}$ (see Equation \ref{inner_eq}). As the accretion radius
moves father from the star the potential energy released by the
accreted material is reduced. This effect can be seen when comparing
the \textit{top} and \textit{bottom panels}, although the effect is
marginal for all but the highest displayed accretion rates. The
\textit{left panels} show accretion streams with an areal coverage of
10\% and the \textit{left panels} 1\%. As the areal coverage reduces
the effective temperature of the accretion hot spot increases,
resulting in an increase in accretion flux, and resulting shift to
bluer wavelengths of the peak flux. This can be seen clearly by
comparing the \textit{left} and \textit{right panels} of Figure
\ref{acc_dom}. Perhaps the most important, albeit qualitative, result
shown is Figure \ref{acc_dom} is an insight into the accretion rate at
which the accretion blackbody flux dominates over the photospheric
flux. Figure \ref{acc_dom} shows that as the accretion rate raises
above \logmdot = $-9$ for systems with 1 or 10\% areal coverage, the
accretion flux dominates the emergent SED at both periods. Therefore
for reasonable coverages (1-10\%) and rotational periods (0.5-5 days)
the photospheric flux is effectively swamped by accretion flux for
accretion rates \logmdot$>-9$. This suggests that accreting BD systems
with accretion rates as low as \logmdot$=-9$ could be difficult to
identify from SEDs as BD stars, and may well be classified as higher
mass stars.

>>>>

\begin{table*}
\begin{tabular}{|l|l|}
\hline
Input parameter&Values (\# values)\\
\hline
Mass ($M_{\odot}$)&0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07 \& 0.08 (8)\\
Age (Myr)&1 \& 10 (2)\\
Rotation period (days)&0.5 \& 5 (2)\\
Areal coverage (of $\dot{M}$, \%)&1 \& 10 (2)\\
Accretion rate (log$(\frac{\dot{M}}{M_{\odot}} yr^{-1})$)&$-6$, $-7$, $-8$, $-9$, $-10$,
$-11$ \& $-12$ (7)\\
Disc mass ($M_*$)&0.01 (1)\\
Surface density profile & $r^{-1}$ (1) \\
Inclination ($^{\circ}$)&0, 27, 39, 48, 56, 64, 71, 77, 84 \& 90
(10)\\
\hline
\end{tabular}
\caption{List of all varied input parameters. Resulting in a total
  number of models of 448 (plus 40 models without radiative transfer
  simulations for the naked BDs) and 4480 SEDs (plus 40 for naked
  BDs). \label{par_space_table}}
\end{table*}

@@

\subsubsection{Inner edge of the dust disc: location}
\label{inner_edge}

@@

As the inner edge moves its temperature is expected to change, this
has been predicted to lead to a correlation of inner edge position and
IR excess \citep{meyer_1997}. This may act to bias surveys correlating
rotation rates with IR excess to search for evidence of
disc-locking. However, the flux from the inner edge will usually peak
between 2 and 3\,$\mu$m \citep{dullemond_2001}, given that the dust
sublimation temperature peaks at $\approx$1400\,K for canonical
densities. This means that for models where dust is significantly
sublimated the inner edge will {\em usually} have a temperature of
$\approx$1400\,K and this correlation of disc position and temperature
will be lost (although see below).

Equation \ref{inner_eq} shows that as the rotation period of the
protostar increases the co-rotation radius decreases. This will result
in, initially, shorter period systems having closer and hotter inner
edges than their longer period counterparts. In addition, if the
accretion rate is increased in these systems the incident flux on the
inner wall will increase leading to a rise in the temperature of the
inner edge.  At some point the dust sublimation temperature may be
reached, leading to a change in radial position of the wall. In
addition, the temperature of the inner edge will then tend to the dust
sublimation temperature.

A further complexity arises when one considers that the density of
material in the disc falls with increasing radius from the star
($\rho(r) \propto r^{\alpha}$), and that the dust sublimation
temperature is density dependent \citep{pollack_1994}.  Therefore, for
systems where the inner edge has been eroded significantly from the
co-rotation radius the temperature of the inner edge will fall
systematically as the inner dust disc radius increases.

In order to examine the effect of dust sublimation on our models we
have measured the radius of the inner edge of the dust disc by
integrating from the centre along the midplane until an optical depth
of 2/3 (at 5500\AA) is reached. Figure \ref{inner_temp} shows the
radial density distribution of the disc.  In this case the the final
inner edge radius and therefore temperature is no longer strongly
correlated with the rotation rate. Figure \ref{inner_temp} shows the
final inner edge location ($R_{\rm inner}$, $R_*$) against the
temperature of the inner wall ($T_{\rm inner}$, K). The \textit{left
  panel} shows the systems designated as typical accretors and the
\textit{right panel} those with extreme accretion rates.  For
\textit{both panels} the systems with rotation periods of 0.5 and 5
days are plotted in red and blue respectively. Those systems where the
change in inner radius was greater than 1$R_*$, and therefore classes
as significantly sublimated, are plotted as crosses.

@@

The \textit{left panel} of Figure \ref{inner_temp} shows that, taken
as a whole, our models with typical accretion rates show a clear
correlation between the temperature at the inner edge and the radius
to this boundary. This agrees with the work of \cite{meyer_1997} where
this correlation is found for there $R_{\rm inner}=$1$-$12$R_*$ and
\logmdot = $-9$ to $-5$, for flat disc models.  \cite{meyer_1997} use
this correlation, and the derivation of IR magnitudes, to predict a
relationship between IR excess and radius to the inner wall. Our data
indicate that this correlation, for typically accreting systems will
translate into a correlation between rotation rate and IR excess. This
could have important implications for studies of disc-locking where
disc presence is examined as a function of rotation rates, provide an
intrinsic bias. In practice however, this correlation is weak (in fact
unobservable) in our data due to the combined effects of the inner
disc wall shape, inclination and flaring effects. This is discussed in
more detail in Section \ref{photometry}.

Figure \ref{inner_temp} shows that for those systems where significant
disc erosion ($\Delta R_{\rm inner} > 1.0 R_*$) has occurred the
resulting temperature of the inner edge is weakly correlated with the
inner edge radius, but, critically, not correlated with rotation rate.
The inner edge temperatures of the remaining systems for the extreme
accretors are slightly anti-correlated with the radius to the inner
edge. For the systems with typical acccretion rates, and longer
periods, Figure \ref{inner_temp} shows there is again a weak
correlation between the radius to, and the temperature of the inner
edge. For the shorter period models with typical accretion rates there
is no clear correlation between the temperature at the inner edge and
radius to this edge.

\subsubsection{Inner edge of the dust disc: shape}
\label{inner_edge_shape}

The initial shape of the inner edge of the dust disc is a vertical
wall coincident with the co-rotation radius. In the previous section
we have shown that models with a negligible accretion rate do not
significantly sublimate the dust, and hence the edge remains
vertical. This inner edge is heated by direct radiation from the
protostar, and its scaleheight increases. Disc material behind the
inner rim is shielded from direct radiation and has a smaller
scaleheight, leading to the `puffed up' inner rim predicted by
\cite{dullenmond_2001}. This effect is illustrated in
Figure~\ref{no_sub_183}, a model in which there is negligible dust
sublimation.

@@

As the flux of the central object increases significant dust
sublimation occurs, leading to a change in the radial position of the
dusty inner edge, but also shaping it: The density drops rapidly away
from the midplane, and since the dust sublimation temperature also
falls the edges of the inner dust disc become curved--this the
mechanism described analytically by \cite{isella_2005}. We illustrate
this effect in Figure~\ref{sub_179_a}, which also shows a scaleheight
decrease behind the curved inner rim, over a significantly larger
radial scale than the previous model.

Finally the most extreme accretor (Figure~\ref{sub_181_a}) shows dust
being destroyed to very large radii ($\sim 0.1$\, AU). A curved rim is
present, without any obvious decrease in scaleheight behind the inner
rim. The distance from the central object is such that the vertical
component of gravity is much diminished, and the disc has a
substantial scaleheight at the inner edge, meaning that it reprocesses
significant protostellar radiation leading to a high near-IR excess.

We note that a similar sequence is apparent across the grid for set
masses. However, the balance of the rotation period, and therefore
inner edge location, and age and areal coverage, therefore flux
levels, leads to changes in the accretion rate at which the dust
sublimation starts. However, in almost all cases the dust sublimation
does not become significant until at least \logmdot=$-9$ (as discussed
above).

@@

The resulting converged disc structures, as discussed in Section
\ref{seds} are then used to create simulated SEDs. In this section we
explore features in these derived SEDs and link them back to the
physical structures within the disc discussed in Section
\ref{disc_struct}. For analysis purposes our SEDs will comprise
contributions from four main luminosity sources. Firstly, the
accretion luminosity ($L_{\rm acc}$) and the intrinsic stellar
luminosity ($L_{*}$), will effectively be combined and provide flux at
the shorter wavelength section of the SED. For our parameter space the
temperature of the photospheres has a range of $~$3000\,K$<T_{\rm
  eff}<$1600\,K, this will comprise flux contributions peaking in the
range, 0.97 to 1.8\,$\mu$m ($\lambda_{\rm peak}\propto \frac{1}{T_{\rm
    acc}}$). The accretion spot temperatures, for typical accretors,
will range from $~$340$>T_{\rm acc}>$9400\,K, emitting in the range
0.3 to 8.5\,$\mu$m. For the extreme accretors temperatures which can
reach as high as 53\,000\,K, provide flux to $~$0.05\,$\mu$m. The
second pair of luminosity sources are from constituent parts of the
circumstellar disc. The inner edge, ($L_{\rm wall}$), at temperatures
of $~$370 to $~$1500\,K, or flux contributions from $~$1.85 to
7.7\,$\mu$m \citep[nominally 2 to 3\,$\mu$m][]{dullemond_2001}, and
the outer disc ($L_{\rm disc}$), at temperatures of $~$300\,K or less
contributing flux longward of $~$10\,$\mu$m. Disentangling the
accretion and stellar photospheroc flux becomes difficult as $T_{\rm
  acc} \to T_{\rm eff}$, which for an acretion rate of -9 \logmdot,
occurs at a fractional coverages of 20 and 10\%, for rotation periods
of 5 and 0.5 days (with $M_*=0.04M_{\odot}$ and an age of 1 Myr). As
discussed in Section \ref{seds} the {\sc torus} radiative transfer
code tags the photon packets before they reach the simulated observer
in one of four ways, stellar and thermal, either direct or
scattered. These components allow us to analyse the contributions from
the main flux contributors discussed, i.e. $L_{\rm acc}$ and $L_*$
equals stellar scattered or direct, with thermal emission comprising
$L_{\rm wall}$ and $L_{\rm disc}$.

Figure \ref{disc_eg_173} shows a typically accreting BDD system with
the component photon contributions highlighted, with
$M_*=0.04M_{\odot}$ at 1 Myr with an areal coverage of 10\%, accretion
rate of \logmdot=$-10$ and rotation period of 0.5 days. The solid
black line is the total simulated SED, with the stellar photons shown
as blue lines and the thermal as red. The scattered and direct
contributions are then shown as dashed and solid lines
respectively. The horizontal and vertical dotted lines then show the
approximate ranges of the flux contributions from the (from left to
right), $L_{\rm acc}$ (blue), $L_*$ (black), $L_{\rm wall}$ (red) and
$L_{\rm disc}$ (black). The vertical dashed lines at the bottom of
Figure \ref{disc_eg_173} show the approximate ranges of the $V$, $I$,
$J$, and $K$ photometric bands we have adopted.

@@

Figure \ref{disc_eg_173} shows that the photosphere and accretion
luminosities dominate the SED at wavelengths shortward of
$~$1$\,\mu$m. After this, for this model, the SED is dominated by
contributions from the disc, at shorter wavelengths, the inner edge
and then at larger wavelengths the outer disc. As shown in Figure
\ref{disc_eg_173} the intrinsic stellar photosphere and accretion
luminosities overlap, in terms of peak wavelength of emission. This
overlap essentially means that the combined stellar flux ($L_*+L_{\rm
  acc}$) will eventually become dominated by the accretion flux, as
one increases the accretion rate. With the features of the intrinsic
stellar photosphere or atmosphere eventually become veiled by the
accretion flux. This will lead to difficulties in using spectral
features to identify BDD systems for wavelengths shorter than $~$1
$\mu$m, and, of course, for wavelengths longer than this the SED
becomes dominated by disc emission. Therefore, once the SED of a BDD
system, at shorter wavelengths, becomes dominated by accretion
luminosity spectral classification of these systems (or at least their
central stars) will become unreliable.

@@

Figure \ref{acc_dom} shows the affect of increasing the accretion
blackbody flux (for increasing accretion rates) for a BD star,
$M=$0.04$M_{\odot}$, at 1 Myr. Whilst the photons originating from the
star (both from $L_*$ and $L_{\rm acc}$) will be tagged as stellar by
{\sc torus} we can separate these flux contributions simply by
observing the naked star system. The panels in Figure \ref{acc_dom}
show the flux from a naked system, with no treatment of the disc. This
enables us to view the effect of increasing accretion rate on the
photospheric flux in isolation. The accretion rates included in all
panels are \logmdot = $-8$, $-9$ and $-12$ (blue, black and red lines
respectively). The \textit{bottom panels} show the systems with a
rotation period of 5 days and \textit{top panels} for those with a
rotation period of 0.5 days. Given our assumption that accretion
occurs from the co-rotation radius, decreasing the rotational period
moves this accretion radius closer to the star, $R_{\rm inner}\propto
\tau^{2/3}$ (see Equation \ref{inner_eq}). As the accretion radius
moves father from the star the potential energy released by the
accreted material is reduced (XXXX WHAT?). This effect can be seen
when comparing the \textit{top} and \textit{bottom panels}, although
the effect is marginal for all but the highest displayed accretion
rates.
\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/acc_dom.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{Figure showing the photospheric flux (log(ergs$/s/cm^2/{\rm
     \AA}$)) against $\lambda$ (log($\mu$m)) of a Brown Dwarf with
   $M=$0.04$M_{\odot}$, at 1 Myr. No disc is included, but blackbody
   fluxes from an accretion stream at the rates of \logmdot $=-8$,
   $-9$ and $-12$ are shown as blue, black and red lines
   respectively. The \textit{bottom panels} show accretion for a star
   rotating at 5 days, with the \textit{top panels} showing that of
   0.5 days. The \textit{left panels} systems with an areal coverage
   of 10\% and the \textit{right panels} has 1\%. The vertical and
   horizontal dashed lines (in the \textit{lower panels}) denote the
   approximate sensitivity ranges of our chosen $V$, $I$, $J$ and $K$
   filters.\label{acc_dom}}
\end{figure*}
